---
title: "depth_native functions"
output: 
  html_document:
    code_folding: hide
runtime: shiny
---


## Select sub, Zthr and Zcontrast
```{r, message=F}

# All of the following will become selector widgets.
# While they are commented here, they belong to the main process below, as 
# now there are only function definition above that.

# SUBID = "sub_02"     # df %>% distinct(sub) to see all available sub
# Zcontrast <- "thresh_zstat1"
# Zthr <- 2.3
# clusterSizeThr <- 50  # to build a meaningful histogram 
# h_breaks <- seq(0, 1, 0.05) # set hist breaks to a fixed range

# -----------------------------------------------------------------------------


library(shiny)
library(stringr)
library(tidyr)
library(dplyr)
library(purrr)
select <- dplyr::select
library(ggplot2)
library(ggthemes)
library(tictoc)


# Immutable parameters
gitdir <- "/data00/layerfMRI/Github_repo/"
bd <- paste0(gitdir,"layerfMRI/analyses/depth_native/")


julabels <- read.csv("labels_juelich.csv") %>% 
  mutate(name = str_replace(name,"['-/]","")) %>%    # get rid of special chars
  mutate(numba = index + 1) %>% 
  select(-index)

```




## Create dictionary of data from `list.files`
The magic is provided by `tidyr::separate`

```{r message=FALSE}

create_dizio_files <- function(data_native_dir = "data_native") {
  
  df <- list.files("data_native", recursive = T) %>% as.data.frame()
  names(df) <- "fname"
  
  
  df <- df %>%
    rowwise() %>%
    tidyr::separate(
      fname, c("sub","contrast","taskrun","zstat"),
      sep = "/", fill = "right", remove = FALSE
    ) %>%
    mutate(taskrun = str_extract(taskrun,"task_[1-4]\\_run_[1-2]")) %>%
    mutate(zstat = str_extract(zstat,"thresh_zstat[1-4]")) %>%
    mutate(pathname = paste0(bd,"/data_native/",fname)) %>%
    dplyr::select(-fname)
  
  return(df)
}



# ------------------------------  Code for Main -------------------------------
# df <- create_dizio_files()


# Display the dataset tree
df <- create_dizio_files()
# show df tree
library(collapsibleTree)
dataset <- df %>% dplyr::select(sub,contrast,taskrun,zstat)
collapsibleTree(dataset, c("sub","contrast","taskrun","zstat"), collapsed = T, zoomable = F)

```



## Load Z, D, JU and extract values for sig voxels
Load the nii for Zmaps (thresh_Zstat[1..4]), Depth and JUelich atlases, all in native space:

- Z_nii  : 8 for each contrast (Motion, Scrambled, M>S, S>M)
- D_nii  : 8, one for each taskrun
- JU_nii : 8, one for each taskrun

Then for each run/contrast
1. threshold the Z_nii to Zthr and create a Zthr_idx of suprathreshold voxels
2. extract Z_vals, D_vals and JU_vals at the locations of Zthr_idx

```{r, message=F}
library(RNifti)

load_niis <- function(SUBID, Zcontrast) {

  # Load the Z_nii for a given contrast (8, one for each taskrun)
  Z_nii <- df %>% 
    filter(sub == SUBID, contrast == "Zstat", zstat == Zcontrast) %>% 
    select(pathname) %>% sapply(readNifti)

  # Load the D_nii (8 taskrun)
  D_nii <- df %>% 
    filter(sub == SUBID, contrast == "depth") %>% 
    select(pathname) %>% sapply(readNifti)

  # Load the JU_nii (8 taskrun)
  JU_nii <- df %>% 
    filter(sub == SUBID, contrast == "atlas") %>% 
    select(pathname) %>% sapply(readNifti)
  
  niis <- list(
    Z = Z_nii, 
    D = D_nii, 
    JU = JU_nii
  )
  
  return(niis)
}


get_idx_and_vals <- function(Zthr, Z_nii, D_nii, JU_nii) {
  
  # Create an index of sig voxels according to Zthr
  Zthr_idx <- sapply(Z_nii, function(x) {which(x > Zthr)})
  
  # Extract Z, D, JU values at the location of Zthr_idx
  Z_vals <- mapply( function(vols,idx) vols[idx], Z_nii, Zthr_idx )
  D_vals <- mapply( function(depth,idx) depth[idx], D_nii, Zthr_idx )
  JU_vals <- mapply( function(atlas,idx) atlas[idx], JU_nii, Zthr_idx )
  
  idx_and_vals <- list(
    idx = Zthr_idx, 
    zvals = Z_vals, 
    depthvals = D_vals, 
    juelichvals = JU_vals
  )
  
  return(idx_and_vals)
}


# ------------------------------  Code for Main -------------------------------

# niis <- load_niis(SUBID, Zcontrast)
# 
# Z_nii  <- niis$Z
# D_nii  <- niis$D
# JU_nii <- niis$JU
# 
# idx_and_vals <- get_idx_and_vals(Zthr, Z_nii, D_nii, JU_nii)
# 
# Zthr_idx <- idx_and_vals$idx  
# Z_vals   <- idx_and_vals$zvals 
# D_vals   <- idx_and_vals$depthvals
# JU_vals  <- idx_and_vals$juelichvals


```




## Hist for each Juelich region
- The (outer) Mapply loops on pairs of `[jv = JU_vals[i], dv = D_vals[i]]` for each i-th taskrun
- The (inner) Sapply loops on juelich regions if the numba sig voxels is > clusterSizeThr
- A final Sapply removes the juelich regions where the counts are NULL


```{r}


get_hist_stats_juelich_ROIs <- function(julabels, JU_vals, D_vals, 
                                        clusterSizeThr, h_breaks, thr_Nrun_sig) {

  # Transform julabels to a list to use sapply
  julabels_list <- spread(julabels, key = name, value = numba) %>% as.list()
    
  # Create a list of length == numba taskruns (8), each one containing 
  # a list of counts for each region where the numba of sig voxels is > clusterSizeThr
  tot <- mapply(function(jv, dv) {
    
    taskrun_count <- sapply(julabels_list, function(ROInumba) {
      
      idxROI <- which(jv == ROInumba)
        
      if (length(idxROI) > clusterSizeThr) {
        hcounts <- hist(dv[idxROI], breaks = h_breaks, plot = F)$counts
      }
    })
    
    taskrun_count <- taskrun_count[!sapply(taskrun_count,is.null)]
    
  }, JU_vals, D_vals) %>% purrr::flatten()  # remove the outer list layer (i.e. taskrun)
  
  
  
  # determine unique sig ROI names
  activeROIs <- tot %>% names() %>% unique()
  
  # group by ROI
  grouped <- map(activeROIs, ~ reduce(tot[names(tot) == .x], rbind) ) %>% setNames(activeROIs)
  
  # (1) remove ROIs where a sig result occurs only in one run
  grouped <- grouped[!sapply(grouped,function(x) is.null(dim(x)) ) ]
  
  # (2) remove ROIs where a sig result occurs in less than 6/8 runs
  grouped <- grouped[!sapply(grouped,function(x) dim(x)[1] < thr_Nrun_sig)]
  
  
  # transform to probabilities, i.e. divide the counts in each row by the sum across columns
  grouped <- lapply(grouped, function(x) x/apply(x, MARGIN = 1, sum))

  # # just checking that the sum of probabilities gives 1
  # # (i.e. to check that I summed across the right margin)
  # lapply(grouped, function(x) apply(x, MARGIN = 1, sum)) %>% print()

    
  # calculate mean p for each depth across taskrun, i.e. across rows
  MUs  <- lapply(grouped, function(x) apply(x, MARGIN = 2, mean)) %>% as.data.frame()
  
  # calculate standard error of p across depth
  sterr <- function(x) {sd(x)/sqrt(length(x))}
  ERRs <- lapply(grouped, function(x) apply(x, MARGIN = 2, sterr)) %>% as.data.frame()
  
  hist_stats = list(
    means = MUs,
    error = ERRs
  )
  
  return(hist_stats)
}


# ------------------------------  Code for Main -------------------------------

# thr_Nrun_sig=6, 
# clusterSizeThr = 50, 
# h_breaks = seq(0, 1, 0.05)

# hist_stats <- get_hist_stats_juelich_ROIs(julabels, JU_vals, D_vals, clusterSizeThr, h_breaks, thr_Nrun_sig)


```


## Main function to get contrast-specific results
This uses all the functions created above, and returns the mean and sterr of counts
across taskrun for a specific contrast.\
I **need** to make this function since I need to get the histograms for _two_ contrasts at a time. Rewriting everything would be prone to mistakes and not elegant (i.e. easy to read/edit).

```
get_contrast_specific_results(Zcontrast)
|
├── load_niis(SUBID, Zcontrast)
├── get_idx_and_vals(Zthr, Z_nii, D_nii, JU_nii)
└── get_hist_stats_juelich_ROIs(julabels, JU_vals, D_vals,
                                clusterSizeThr, h_breaks, thr_Nrun_sig)

```

```{r, message=FALSE}

# Main function, to be run for each contrast
get_contrast_specific_results <- function(SUBID, Zcontrast, Zthr, julabels, 
                                          JU_vals, D_vals, clusterSizeThr, h_breaks, thr_Nrun_sig) {
  
  # Load Z, D, JU niftis (24)
  niis <- load_niis(SUBID, Zcontrast)
  
  Z_nii  <- niis$Z
  D_nii  <- niis$D
  JU_nii <- niis$JU
  
  
  # Extract values for sig voxels in Zstat, Depth and JUelich native maps
  idx_and_vals <- get_idx_and_vals(Zthr, Z_nii, D_nii, JU_nii)
  
  Zthr_idx <- idx_and_vals$idx  
  Z_vals   <- idx_and_vals$zvals 
  D_vals   <- idx_and_vals$depthvals
  JU_vals  <- idx_and_vals$juelichvals
  
  
  # (1) Estimate the distribution of sig voxels at different depths for each ROI in the juelich atlas.
  # (2) Exclude ROIs where the numba of sig voxels is < clusterSizeThr in less than 6/8 runs
  # (3) Retain mean and standard error across runs
  hist_stats <- get_hist_stats_juelich_ROIs(julabels, JU_vals, D_vals,
                                            clusterSizeThr, h_breaks, thr_Nrun_sig)
  
  return(hist_stats)
  
}



```


## Prepare UI
```{r, echo=F}

df <- create_dizio_files()

selectInput("chosen_sub", label = "Select participant", choices = df %>% distinct(sub))
sliderInput("chosen_zthr", label = "Z threshold", min = 2.3, max = 3.5, value = 2.3)
sliderInput("chosen_clusterSizeThr", label = "min cluster size", min = 50, max = 300, value = 50)
sliderInput("chosen_nbins", "bins for sampling cortical depth", min = 10, max = 30, value = 20)
sliderInput("chosen_thr_Nrun_sig", "min numba of run with min cluster size", min = 2, max = 8, value = 4)



plotOutput("resPlot", width = "100%", height = "600px")  





```



## Main process
```{r, fig.width=10, fig.height=10}

# df %>% dplyr::select(sub) %>% unique()

# Create dictionary of data from list.files()
df <- create_dizio_files()

results <- reactive({
  
  req( c(input$chosen_sub, input$chosen_zthr, input$chosen_clusterSizeThr, 
         input$chosen_nbins, input$chosen_thr_Nrun_sig ) )
  
  SUBID = input$chosen_sub     # df %>% distinct(sub) to see all available sub
  
  Zthr <- input$chosen_zthr
  
  clusterSizeThr <- input$chosen_clusterSizeThr        # to build a meaningful histogram
  
  binwidth <- 1/input$chosen_nbins
  h_breaks <- seq(0, 1, binwidth)          # set hist breaks to a fixed range
  
  thr_Nrun_sig <- input$chosen_thr_Nrun_sig  # min numba of run with min cluster size
  
  
  Motion_res <- get_contrast_specific_results(SUBID, Zcontrast = "thresh_zstat1", Zthr, julabels, 
                                              JU_vals, D_vals, clusterSizeThr, h_breaks, thr_Nrun_sig)

  Scrambled_res <- get_contrast_specific_results(SUBID, Zcontrast = "thresh_zstat2", Zthr, julabels, 
                                              JU_vals, D_vals, clusterSizeThr, h_breaks, thr_Nrun_sig)
  
  list(M = Motion_res, S = Scrambled_res)
})


observe({
  
  req(results())
  req(input$chosen_nbins)
  
  binwidth <- 1/input$chosen_nbins
  
  Mot_res <- results()$M
  Scr_res <- results()$S
  
  # Produce a plot only if there is still something sig at the given thresholds
  numba_plots <- length(Mot_res$means) + length(Scr_res$means)
  
  if (numba_plots == 0) {nullvar = NULL} else {nullvar = "ce sta"}
  req(nullvar)
  
  observeEvent(
    nullvar,
    {
    # Prepare the results for plotting using the magic tidyverse:
    # (1) put the resulting df for all contrasts in one list
    # (2) add depth and contrast factor
    # (3) tidy (i.e. WIDE -> LONG)
    # (4) bind_rows all the df into one
    res_means <- list(Mot_res$means, Scr_res$means) %>%
      map(~ .x %>% mutate(depth = 1:nrow(.x)/nrow(.x)) ) %>%
      map2(. , c("Motion","Scrambled"),  ~ .x %>% mutate(contrast = .y)) %>%
      map(~ .x %>% gather(contains("GM"), key = "ROI", value = "proportion") ) %>%
      reduce(bind_rows)

    res_error <- list(Mot_res$error, Scr_res$error) %>%
      map(~ .x %>% mutate(depth = 1:nrow(.x)/nrow(.x)) ) %>%
      map2(. , c("Motion","Scrambled"),  ~ .x %>% mutate(contrast = .y)) %>%
      map(~ .x %>% gather(contains("GM"), key = "ROI", value = "error") ) %>%
      reduce(bind_rows)

    res <- inner_join(res_means, res_error, c("contrast","ROI","depth"))
    
    # prepare and render the image to plot
    fig <- res %>%
      mutate(ROI = str_replace_all(ROI, "_", " ")) %>%
      ggplot( aes(x = depth, y = proportion, fill = contrast) ) +
      geom_bar(stat = "identity", width = binwidth, position = position_identity(), alpha = 0.5) +
      geom_linerange(aes(ymin = proportion - error, ymax = proportion + error, color=contrast, alpha = 0.7)) +
      facet_wrap( ~ ROI, scales = "free", labeller = label_wrap_gen(width=22)) +
      theme_minimal()

    
    output$resPlot <- renderPlot(fig)
    
    }
  )
  
  
})


  














```



## Smooth lines with two contrasts and errorbars
```{r, fig.width=10, fig.height=6, message=F}
# library(ggthemes)
# 
# res %>%
#   mutate(ROI = str_replace_all(ROI, "_", " ")) %>% 
#   ggplot( aes(x = depth, y = proportion, color = contrast) ) +
#   geom_smooth(alpha = 0.5, se=F, n=20, na.rm = T, method = "loess") +
#   geom_linerange(aes(ymin = proportion - error, ymax = proportion + error, color=contrast, alpha = 0.7)) +
#   facet_wrap( ~ ROI, scales = "free", labeller = label_wrap_gen(width=22)) +
#   theme_minimal()
# 


```












