---
title: "Fixing the presentation log file"
author: "Leonardo Cerliani"
date: '`r paste("Created on March 2, 2021. Last Update on", Sys.Date())`'
output: 
  html_document:
    code_folding: hide
---

## The problem: movies with `end_TR` logged **after** the end of the fmri4D acquisition

The `end_TR` of some movies goes beyond the number of volumes in the actual fMRI4D. This does not happen in all cases.

This problem is sneaky since the `end_TR` is always consistent with the `total_TR` in the log file, however there is a discrepancy between the `total_TR` and the actual number of volumes in the fMRI4D nifti file (as read from the actual nifti files).

As a result, the code which should extract the time courses for each movie by looking at the `start_TR` and `end_TR` sometimes breaks with a typical `index exceeds dimensions`.

```
movie_time_course = fmri4D[idx_voxels, start_TR : (end_TR - 1)]
```

Ritu reports that this is due to the fact that the presentation log kept recording after the stimuli presentation was terminated, into the following sequence, which is a 5 volumes top-up sequence. Indeed the difference between the numba of volumes and the `total_TR` in the log - when there is one - is always 5 volumes.

I will fix in the following way: 

1. After reading the original `log_summary.csv` I will use `fslr` to read the header of each fMRI4D file, extract the numba of volumes  and send it to a new column `dim4`. 
2. I will then check `ifelse` the `end_TR` of any video is bigger than `dim4`, and in this case set it to `dim4`.
3. Finally I will write a new `log_summary_FIXED.csv` which will be used for the time courses extraction.

```{r, message=F, warning=F}
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(tictoc)
library(kableExtra)
library(DT)

library(fslr)
get.fsloutput()
get.fsl()

# Immutable parameters
gitdir <- "/data00/layerfMRI/Github_repo/"
bd <- paste0(gitdir,"layerfMRI/analyses/dual_ISC/")
regdatadir <- "/data00/layerfMRI/regdata/"
depthdatadir <- paste0(bd, "/data_native/")


df <- read.csv("log_summary.csv", stringsAsFactors = F) %>% 
  arrange(subject,session,task,run) %>% 
  select(-c(Trial,Event.Type,Time,NMov)) %>% 
  mutate(sub = sprintf("sub_%02d",subject) ) %>%
  select(-subject) %>% 
  mutate(ses = sprintf("ses_%02d", session) ) %>% 
  select(-session) %>% 
  mutate(contrast = ifelse(Type == "M", "Motion", "Scrambled") ) %>% 
  mutate(contrast_file = ifelse(Type == "M", "thresh_zstat1.nii.gz", "thresh_zstat2.nii.gz") ) %>% 
  select(-c(Type) ) %>% 
  rename(muvi = Title) %>% 
  relocate(sub, ses, task, run, muvi, start_TR, end_TR, contrast, contrast_file)


# # just check the differences between Duration and ExpectedDuration
# df %>% 
#   rowwise() %>% 
#   mutate(diffDuration = ifelse(Duration != ExpectedDuration, "poppa",NA)) %>% 
#   filter(!is.na(diffDuration))

```



## 1. Display the issue
You can see the problem in the table below, because `dim4` is higher than `end_TR` in some movies.
NB: `sub_08` has been removed.
```{r}
# function to get the dim4 of the fmri4D without loading them - to be passed to the map below
get_fmri4D_dim4 <- function(df_sub_taskrun) {
  this <- df_sub_taskrun[1,] %>% as.list()
  fmri4D_file <- paste0(regdatadir,"/",this$sub,"/", 
                        this$ses,"/func/task_",this$task,"_run_", this$run,"_4D.nii.gz")
  dim4_info_raw <- fslr::fslhd(fmri4D_file, verbose = F)[9]
  dim4_info <- str_split(dim4_info_raw,"\t")[[1]] %>% tail(1) %>% as.numeric()
  df_sub_taskrun %>% mutate(dim4 = dim4_info)
}

# get the dim4 as a new column
pf <- df %>% 
  group_by(sub,task,run) %>% 
  group_split() %>% 
  map(~ .x %>% get_fmri4D_dim4) %>% 
  bind_rows()


# you can see the problem here, because the dim4 is higher than the end_TR in some movies
pf %>% 
  mutate(dim4 = as.numeric(dim4)) %>% 
  filter(end_TR > dim4) %>% 
  filter(sub != "sub_08") %>%
  mutate(diff_end_dim4 = dim4 - end_TR) %>% 
  relocate(dim4, .after = total_TR) %>% 
  relocate(Duration, .after = muvi) %>% 
  select(-c(contrast,contrast_file)) %>% 
  datatable() %>% 
  formatStyle(
    c("end_TR","total_TR","dim4"), backgroundColor = "#F3F6A2"
  )





```



## 2. Fixing the issue

```{r}

# fixing the issue

pf %>%
  filter(sub != "sub_08") %>%
  mutate(end_TR_new = ifelse(end_TR > dim4, dim4, end_TR) ) %>%  # here's the fix
  relocate(end_TR_new, .after=end_TR) %>%
  filter(end_TR > dim4) %>%
  select(-c(contrast, contrast_file,Duration)) %>% 
  datatable() %>% 
  formatStyle(
    c("end_TR","total_TR","dim4"), backgroundColor = "#F3F6A2"
  )


```


## 3. Saving a new `log_summary_FIXED.csv` file
And since I'm in at, I will also filter out `sub_08`

```{r}

df_FIXED <- pf %>%
  filter(sub != "sub_08") %>%
  mutate(end_TR_new = ifelse(end_TR > dim4, dim4, end_TR) ) %>%  # here's the fix
  relocate(end_TR_new, .after=end_TR) %>%
  mutate(end_TR = end_TR_new) %>% 
  # mutate(diffTR = end_TR - end_TR_new) # just to check that is 0 everywhere
  # mutate(PROBLEM = ifelse( (dim4 - end_TR) < 0, "PROBLEM", NA ) ) %>%  another check
  select(-end_TR_new) %>%   # now we can remove it and leave only end_TR
  select(-c(total_TR, Duration) )  # rm as inconsistent with dim4 thus confusing
  
  
df_FIXED %>% write.csv("log_summary_FIXED.csv", row.names = F)


read.csv("log_summary_FIXED.csv", stringsAsFactors = F) %>% datatable()

```




























