---
title: "elements 2D histograms"
output: html_notebook
runtime: shiny
---

```{r, librerie, message=FALSE, echo=FALSE}
library(shiny)
library(plotly)
library(oro.nifti)
library(neurobase)
library(papayaWidget)
library(magrittr)

bd <- "/data00/layerfMRI/Github_repo/layerfMRI/analyses/2D_histogram"
datadir <- paste0(bd,"/data4histograms")
ROIsdir <- paste0(bd,"/ROIs")

MNI <- paste0(datadir,"/icbm152_2009_brain.nii.gz") %>% oro.nifti::readNIfTI()

Zmotion <- paste0(datadir,"/MOTION_thresh_zstat1.nii.gz") %>% oro.nifti::readNIfTI()
Zscrambled <- paste0(datadir,"/SCRAMBLED_thresh_zstat1.nii.gz") %>% oro.nifti::readNIfTI()

depth <- paste0(datadir,"/median_depth.nii.gz") %>% oro.nifti::readNIfTI()
layers <- paste0(datadir,"/median_layers.nii.gz") %>% oro.nifti::readNIfTI()

```



# One dimensional histogram as a first quick check
```{r, 1D histogram, echo=FALSE}

# # First attempt (not interactive) with bars
# plot_ly(alpha = 0.6) %>%
#   add_histogram(x = depth[idx_motion]) %>%
#   add_histogram(x = depth[idx_scrambled]) %>%
#   layout(barmode = "overlay")


# Zthr <-  2.3

plot_1Dhist <- function(Zthr) {

  idx_motion <- which(Zmotion > Zthr)
  idx_scrambled <- which(Zscrambled > Zthr)
  
  M_density <- density(depth[idx_motion])
  S_density <- density(depth[idx_scrambled])

  p <- plot_ly() %>% 
    add_trace(
      x = M_density$x, y = M_density$y, 
      type = 'scatter', mode = 'lines', name = 'Motion', fill = 'tozeroy'    
    ) %>% 
    add_trace(
      x = S_density$x, y = S_density$y, 
      type = 'scatter', mode = 'lines', name = 'Scrambled', fill = 'tozeroy'
    )
  
  return(p)
  
}


# ------------ UI -------------
renderPlotly({
  plot_1Dhist(input$Zthr)  
})

sliderInput("Zthr", label = "Zthr", min = 2.3, max = 8, value = 3.1)


```




# 2D histogram

The easiest and quickest way to create a 2D histogram is to pass, for every voxel (above a certain Z threshold) (1) its Z-value and (2) its cortical depth value. This will allow creating a scatterplot from which the 2D histogram can be estimated in plotly using the `add_histogram2dcontour` trace.

If instead you wish to fit a bivariate distribution, follow [this link](https://www.datacamp.com/community/tutorials/bivariate-heatmaps)


```{r 2D histograms, echo=FALSE}

# Zthr <-  2.3
# Zmax <- 8
# chosenZstat <- Zmotion
# layeringmap <- depth   # layers or depth

# # layeringmap is layers or depth

plot_2Dhist <- function(Zthr, Zmax, chosenZstat, title) {

  idx <- which(chosenZstat > Zthr & chosenZstat < Zmax)
  x = depth[idx]
  y = chosenZstat[idx]
  
  p <- plot_ly(
    x = depth[idx],
    y = chosenZstat[idx]
  ) %>% 
    add_markers(color = x, colors = "Reds", alpha = 0) %>%
    add_trace(
      type = 'histogram2dcontour',
      colorscale = "Viridis",
      contours = list(showlabels=TRUE)
    ) %>% 
    layout(
      title = title,
      xaxis = list(
        title = 'cortical depth'
      ),
      yaxis = list(
        title = 'Z-stat'
      )
    )
  
  return(p)

}


# ------------ UI -------------
renderPlotly({
  plot_2Dhist(
    Zthr = input$Zthr2D, Zmax = 8,
    chosenZstat = Zmotion, title = 'Motion'
  )
})

sliderInput("Zthr2D", label = "Zthr", min = 2.3, max = 8, value = 3.1)

renderPlotly({
  plot_2Dhist(
    Zthr = input$Zthr2D, Zmax = 8,
    chosenZstat = Zscrambled, title = 'Scrambled'
  )
})



```



# kernel density estimate with `MASS:kde2`
```{r, kernel density estimate}
library(MASS)

thr <- 5

idx <- which(Zmotion > thr & Zmotion < 8)
dM <- kde2d(Zmotion[idx], depth[idx], n = 100)
renderPlot(contour(dM)) 


idx <- which(Zscrambled > thr & Zscrambled < 8)
dS <- kde2d(Zscrambled[idx], depth[idx], n = 100)
renderPlot(contour(dS))


```







# Brain viewer heatmap
```{r, heatmap brain viewer, echo=FALSE}


# # ---------- functions -------------
# plot_Zstat <- function(slice, chosenZstat, Zthr) {
# 
#   Zstat <- chosenZstat
# 
#   Zstat[Zstat < Zthr] <- NA
#   Zstat[Zstat >= Zthr] <- 1
#   
#   MNI[MNI < 50] <- NA
# 
#   p <- plot_ly(width = 280, height = 300) %>%
#     add_trace(
#       z = MNI[ , ,input$slice], transpose = TRUE,
#       type = "heatmap",
#       colorscale = "Greys", showscale = FALSE, opacity = 1
#     ) %>%
#     add_trace(
#       z = Zstat[ , ,input$slice], transpose = TRUE,
#       type = "heatmap",
#       colorscale = "Reds", showscale = FALSE, opacity = 1
#     )
# 
# }
# 
# # ------------------ UI ---------------------------
# 
# output$plotSlice <-  renderPlotly({
#   plot_Zstat(input$slice, chosenZstat = Zmotion, Zthr = input$Zthr)
# })
# 
# sliderInput("slice", label = "slice", min = 50, max = 150, value = 100)
# 
# plotlyOutput("plotSlice")



```



















# Papaya viewer
Maybe not the best: it is a bit slow to load the images. On the other hand the threshold of images to display could probably be passed reactively <- no it doesn't work. Ok no papaya viewer, we go with heatmaps
```{r, echo=FALSE}

# sliderInput("Zthr", label = "Z threshold", min = 1.9, max = 5, value = 2.3)
# 
# radioButtons(
#   inputId = "ZstatChosen", label = "Choose task",
#   choices = list("Motion" = Zmotion, "Scrambled" = Zscrambled)
# )
# 
# papayaOutput(outputId = 'viewer', height = "600px")
# 
# output$viewer <- renderPapaya({
#   papaya(
#     img = c(MNI,input$ZstatChosen),
#     options = list(
#       papayaOptions(min = 50, max = 100),
#       papayaOptions(min = input$Zstat)
#     )
#   )
# })


```




