---
title: "Checks on muvi duration in logfiles"
output: html_document
---

While preparing concatenated time courses of all muvis for a Tensor ICA (TICA) analysis, I noticed that the total length was different across subjects. This should prevent the possibility of running not only TICA, but also ISC.

I decided to investigate further. I previously calculated the length of each muvi as `end_TR - start_TR`, so I first decided to instead check if the total lenght would be the same if, instead of using the _calculated_ length, I would use the _expected_ length of the muvi (which is also reported in the log file)


## Load libraries + csv of log_summary
NB: The `log_summary_FIXED.csv` contains the same info of the `log_summary.csv` 
prepared by Lorenzo, with just some additional columns and some fixes in the values -
e.g. "sub_02" instead of "2" in the sub column - to fit my processing scripts

```{r, message=F}
library(tidyverse)
library(DT)
library(kableExtra)
library(rmarkdown)

df <- read_csv("log_summary_FIXED.csv")

```



## Checks on the _expected_ duration
Check that all the concatenated time courses will be of identical length across
subs. This might sound lame - since we use the ExpectedDuration which is a 
constant - however the ExpectedDuration is taken from the log file, and we want
to check that the its value is the same in all subs.
```{r}

# start/end TR and expected duration for each muvi
df %>% 
  group_by(muvi) %>% 
  select(sub, run, muvi, start_TR, end_TR, ExpectedDuration) %>%
  arrange(muvi,run) %>% # group_split() 
  paged_table
  


# In order for ISFC/TICA to be feasible, the concatenated time course
# of all movies must have the same length in all subs
df %>% 
  group_by(sub) %>%
  select(sub,muvi,ExpectedDuration) %>%
  summarise(tot_TR = sum(ExpectedDuration), .groups="drop")


# Checking that a given movie covers the same numba of time points
# in all subs. This is important since the the value could potentially
# differ in the log file of different subs, even though their sum - checked
# above - might still be the same.
# Briefly, the var_nTR in the table below should always be 0
df %>% 
  group_by(muvi) %>%
  summarise(
    nTR_min = min(ExpectedDuration),
    nTR_max = max(ExpectedDuration),
    var_nTR = var(ExpectedDuration),
    .groups = "drop"
  ) %>% 
  paged_table


# # Checking that the end_TR does not go above the recorded length of the fmri 4D
# # (data to be gathered using fsl in the terminal - CORRECT and long to display)
# df %>% 
#   group_by(ses,task,run,sub) %>% 
#   summarise(
#     max_end_TR = max(end_TR),
#     var_end_TR = var(max(end_TR)),
#     .groups = "drop"
#   ) %>%
#   group_by(sub) %>% 
#   group_split()

```



## Checks on the _recorded_ duration - i.e. that recorded in the log file. Problems here.
```{r}
# -------- PROBLEM HERE!! --------------------------------

# Calculate the total lenght of the time course for each sub if we consider
# end_TR - start_TR
df %>% 
  group_by(sub) %>% 
  select(sub,start_TR, end_TR) %>% 
  mutate(CalculatedDuration = end_TR - start_TR) %>% 
  summarise(total_tc_length = sum(CalculatedDuration))


```


The difference between expected and calculated length of the concatenated
time course is not that big (max 7 TR), however there are larger differences
in the calculated vs. expected muvi duration (up to 11 TRs).

This means that while the total length is not very different across subs, the
recorded alignment of the muvis across subs can vary substantially

```{r}
# Mismatch between expected and recorded duration 
# of the muvis.
# Note that there is some similarity across subjects (try group_by(sub) )
# and it happens only for 8 specific intact (M) muvis (try group_by(muvi) )
df %>% 
  group_by(muvi) %>% 
  select(muvi,start_TR, end_TR, ExpectedDuration) %>% 
  mutate(
    CalculatedDuration = end_TR - start_TR,
    Expected_minus_Calculated = ExpectedDuration - CalculatedDuration
  ) %>% 
  filter(Expected_minus_Calculated != 0) %>% 
  # group_split()
  summarise(
    avgMismatch = mean(Expected_minus_Calculated)
  )
  


# However in reality the misalignment does not look that big
n_muvis <- df$muvi %>% unique() %>% length
colori = rainbow(n_muvis, s=.6, v=.9)[sample(1:n_muvis,n_muvis)]

df %>%
  mutate(duration = end_TR - start_TR) %>% 
  select(sub, muvi, duration) %>% 
  arrange(sub,muvi) %>%
  ggplot(aes(fill = muvi, y = duration, x = sub)) +
    geom_bar(position = "stack", stat = "identity") +
    scale_fill_manual(values=colori) +
    theme(axis.text.x=element_text(angle=45, hjust=1))

```




## Fixing the problem using the expected rather than the recorded duration of the movie
```{r}
# To fix this, we can use the expected duration instead of the calculated. 
# As we saw above, this will obviously allow to have concatenated time
# courses of the same lenght
df %>% 
  group_by(sub) %>%
  select(sub,muvi,ExpectedDuration) %>%
  summarise(tot_TR = sum(ExpectedDuration), .groups="drop")


# However I need to make sure that using start_TR + ExpectedDuration
# would not go over the recorded length of the time course.
# Let's check that.
df %>% 
  select(muvi,start_TR, end_TR, ExpectedDuration, dim4) %>% 
  mutate(
    ImposedDuration = start_TR + ExpectedDuration,
    overflow = ImposedDuration - dim4
  ) %>% 
  filter(overflow > 0)


# If I use start_TR + ExpectedDuration, I have three muvis that go
# over the length of the fmri 4D. This is probably because Lorenzo
# calculated the start_TR one volume after the recorded start volume
# of the muvi (it needs a shift -1).
# We can fix this by using start_TR - 1. This is better than
# ExpectedDuration - 1 since it also fixes the issue of the lacked shift.
df %>% 
  select(muvi,start_TR, end_TR, ExpectedDuration, dim4) %>% 
  mutate(
    ImposedDuration = (start_TR - 1) + ExpectedDuration,
    overflow = ImposedDuration - dim4
  ) %>% 
  filter(overflow == 0)


# At this point everything is perfectly aligned
# NB: the colori is defined above, to have the same colors

df %>%
  select(sub, muvi, ExpectedDuration) %>% 
  arrange(sub,muvi) %>%
  ggplot(aes(fill = muvi, y = ExpectedDuration, x = sub)) +
    geom_bar(position = "stack", stat = "identity") +
    scale_fill_manual(values=colori) +
    theme(axis.text.x=element_text(angle=45, hjust=1))


```
































