---
title: "fig. 3 - results ISC2"
subtitle: "Results of the comparison between Scrambled and Motion from the second ISC"
output: 
  html_document:
    code_folding: hide
    keep_md: true
---

## Analytic strategy
We carried out a dual Inter-Subject Correlation (ISC) for each cortical depth bin and Juelich regions (JU ROIs) across all subjects.

**The first ISC** was carried out in the MNI space, after smoothing the data with a 6mm FWHM gaussian kernel (8 times the original voxel size). This ISC was intended to provide localizers of brain activity in either conditions. To this aim, the union of Intact ("Motion", denoted with `M`) and Scrambled (`S`) thresholded Zmaps (`M_OR_S`) was used.

Our prior hypothesis involved considering BA44, PFt and latOCC (lateral occipital). Still in the following we show the results for all the regions in the [Juelich atlas](https://www.fz-juelich.de/inm/inm-1/EN/Forschung/_docs/SPMAnatomyToolbox/SPMAnatomyToolbox_node.html) where at least 100 (native space) voxels (`clusterSizeThr` below) were significantly activated in the `M_OR_S` contrast (`N_sig` below). This assessment was carried out to confirm the specificity of our experimental hypothesis with respect to our regions of interest.


**The second ISC** took as input the average (mean) time courses extracted from the **_unsmoothed_** fmri data with the following level of granularity:

- for each cortical depth bin (user-defined `nbin` below)
- for each JU ROI (N out of ~100 JU ROIs after `clusterSizeThr`)
- for each movie (18 Motion and 18 Scrambled)
- for each run (n_run = 2)
- for each subject (n_subject = 9)

For each cortical depth bin and each JU ROI, we concatenated the segments of the BOLD time course containing movie stimuli, transformed the intensity values in Z-scores and averaged the resulting values across voxels. This resulted in a time course containing the brain activity elicited by the movies for each cortical depth bin, each region of interest and each subject.



## Read csv of ISC results and prepare the environment
The input are the ISC parameter estimates obtained for each bin of each JU region in each subject.


```{r load-libs, message=F}
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(ggplot2)
library(ggthemes)
library(kableExtra)
library(DT)
options(digits = 3)

# -------------------------- User-defined parameters --------------------------

nbin <- 6     # 3-10; check also the filenames in this dir 

# -------------------------- End of User-defined parameters -------------------


# read the table with the names of JU ROI and remove WM regions
julabels <- read.csv("labels_juelich.csv", stringsAsFactors = F) %>% 
  mutate(name = str_replace(name,"['-/]","")) %>%    # get rid of special chars
  mutate(numba = index + 1) %>%
  filter(grepl("GM", name)) %>% 
  select(-index)


results_filename <- paste0(
  getwd(), 
  "/",
  "M_OR_S_JU_time_courses_thr100_bin",
  nbin,
  "_isc.csv"
)

ISC_estimates <- read.csv(results_filename,stringsAsFactors = F)

```



## Data used to generate the plots for **`r nbin` bins**

Current binning: **`r nbin` cortical depth bins**

User-defined parameters in the cell above can be modified to display data and results related to other binning choices.

```{r, message=F}
# Show table with the ISC parameter estimates
ISC_estimates %>%
  select(-c(X,tc_concatenated)) %>%
  inner_join(julabels, by=c("JU"="numba")) %>%
  rename(JU_region = name) %>% 
  select(-JU) %>% 
  relocate(JU_region) %>% 
  datatable(
    extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('csv', 'excel')
  )
  ) %>% 
  formatRound(columns=c('isc'), digits=3)


```







## Test Motion > Scrambled
Separately for each bin of each JU region

```{r t-test, message=F}

ISC_estimates_nest <- ISC_estimates %>%
  select(-c(tc_concatenated, X)) %>%
  inner_join(julabels, by=c("JU"="numba")) %>%
  group_by(JU,D_bins) %>%
  nest()


# Function to test Motion > Scrambled in every nested df
# It's defined outside so that it can be easily modified to be something
# other than t.test
compare_contrasts <- function(df) {
  t.test(df %>% filter(contrast == "Motion") %>% select(isc) %>% pull(), 
         df %>% filter(contrast == "Scrambled") %>% select(isc) %>% pull(), 
         paired = TRUE)
}


# Carry out the comparison Motion > Scrambled for each D_bin in each JU ROI
ttest_res <- ISC_estimates_nest %>% 
  mutate(
    ttest = map(data, compare_contrasts)
  )


library(broom)

# # Use the line below to check which estimates to extract
# ttest_res$ttest[[1]] %>% glance()

# Extract the parameters of interest (T,p)
ttest_res <- ttest_res %>% 
  mutate(
    glance = ttest %>% map(glance),
    T = glance %>% map_dbl("statistic"),
    p = glance %>% map_dbl("p.value")
  )


```



## Results table


```{r create-table}
# Add column of names for JU ROIs
res <- ttest_res %>% 
  select(D_bins, JU, T, p, data) %>% 
  inner_join(. , julabels, by = c("JU"="numba")) %>% 
  mutate(name = str_replace_all(name, "_", " ")) 


# Apply fdr correction per ROI
res <- res %>% 
  group_by(JU) %>%
  mutate(
    pcorr = p.adjust(p, "fdr"),
  ) %>% 
  mutate(pcorr_sig = ifelse(pcorr <= 0.05, D_bins, NA))  # to print asterisk on sig p bins


# Show a table of sig ROIs/bins
title = "Significant differences uncorrected and corrected ( $q(FDR) = 0.05$ ) in Motion > Scrambled"

res %>%
  select(-data) %>% 
  ungroup() %>% 
  filter(!is.na(pcorr_sig)) %>%
  select(-c(JU,pcorr_sig)) %>% 
  relocate(pcorr, .after = p) %>%
  kbl(caption =  title) %>% 
  kable_styling(c("condensed")) 


```




## Plot T stat

```{r plot-group-comparison-T, fig.width=10, fig.height=10}

# Plot the effect size (actually T stat) for the comparison Motion > Scrambled
res %>% 
  mutate(D_bins = as.numeric(D_bins)) %>% 
  ggplot( aes(x = D_bins, y = T, fill = D_bins ) ) +
  # scale_fill_gradient(low = "orange", high = "blue", na.value = NA) +
  scale_fill_gradient2_tableau() +
  geom_bar(stat = "identity", position = position_identity()) +
  geom_text( aes(x = pcorr_sig, y = T + 0.1, label = "*"), fontface = "bold", size=10, na.rm = T ) +
  facet_wrap(~ name, scales = "free", labeller = label_wrap_gen(width=22)) +
  # coord_cartesian(ylim = c(min(res$T), max(res$T) + 0.2)) +
  theme_minimal() +  # also theme_bw and theme_few
  labs(
    title = "T statistic at each cortical depth for Motion > Scrambled",
    subtitle = "Comparison of the ISC parameter estimates in each bin in either contrast"
  ) +
  ylab("Scrambled <   > Motion") + xlab("Cortical Depth bins :  1 = WM/GM border, 6 = pial")+
  theme(
    strip.text.x = element_text(size = 12), # increase size of title for each subplot
    panel.spacing.y = unit(3, "lines")      # increase space between facets
  ) + 
  coord_flip(ylim = c(min(res$T) - 0.2, max(res$T) + 0.2))

```



## Plot descriptives

```{r plot-group-comparison-descriptives, fig.width=10, fig.height=8}

# function to calculate standard error after removing NAs
sterr <- function(x) {
  x <- x[!is.na(x)]
  sd(x)/sqrt(length(x))
}

res_descriptives <- res %>%
  select(-c(T,p,pcorr,name)) %>% 
  unnest(data) %>% 
  mutate(name = str_replace_all(name, "_", " ")) %>% 
  group_by(JU, name, D_bins, contrast, pcorr_sig) %>% 
  summarise(
    meanZ = mean(isc, na.rm = T),
    sterrZ = sterr(isc),
    .groups = "drop"
  )



res_descriptives %>%
  mutate(pcorr_sig = ifelse(contrast == "Motion", pcorr_sig, NA)) %>%  # to plot only one asterisk
  ggplot( aes(x = D_bins, y = meanZ, fill = contrast) ) +
  geom_bar(stat = "identity", position = position_identity(), alpha = 0.6) +
  geom_text(
    aes(x = pcorr_sig, y = meanZ + 0, label = "*"), fontface = "bold", size=10, na.rm = T
  ) +
  geom_linerange(aes(ymin = meanZ - sterrZ, ymax = meanZ + sterrZ, color=contrast), alpha = 0.5) +
  # coord_cartesian(ylim = c(zthr, max(res_descriptives$meanZ) + 0.2)) +
  facet_wrap(~ name, scales = "free", labeller = label_wrap_gen(width=22)) +
  theme_minimal() +
  labs(
    title = "Mean ISC parameter estimates for each Depth bin in each Juelich region across participants"
  ) +
  ylab("Mean ISC estimate across participants") + 
  xlab("Cortical Depth bins :  1 = WM/GM border, 6 = pial") +
  theme(
    strip.text.x = element_text(size = 11), # increase size of title for each subplot
    panel.spacing.y = unit(3, "lines")      # increase space between facets
  ) +
  coord_flip()



```



## Plot T stat with ISC(Motion - Scrambled) variability within subject

```{r plot-group-variability, fig.width=10, fig.height=12}

# function to calculate the mean of Motion - Scrambled across participants
get_M_minus_S <- function(df) {
  df %>% 
    pivot_wider(names_from = contrast, values_from = isc) %>% 
    mutate(M_minus_S = Motion - Scrambled) %>% 
    select(M_minus_S)
}


# function to calculate standard error after removing NAs
sterr <- function(x) {
  x <- x[!is.na(x)]
  sd(x)/sqrt(length(x))
}


# prepare the Motion - Scrambled for each sub, its mean and sderr
res2plot <- res %>% 
  mutate(
    diff_MS = map(data, get_M_minus_S),
  ) %>% 
  select(D_bins, JU, diff_MS,name,pcorr_sig) %>% 
  unnest(diff_MS) %>% 
  group_by(D_bins, JU) %>% 
  mutate(
    diff_mean = mean(M_minus_S, na.rm = T),
    diff_sterr = sterr(M_minus_S)
  )
  # # pipe the following to have a clearer view of res2plot
  # group_by(D_bins, JU, diff_mean,diff_sterr,name,pcorr_sig) %>% 
  # nest()



# plot mean difference Motion - Scrambled, and its variability across subs
res2plot %>% 
  # unnest(data) %>% 
  ggplot( aes(x = D_bins, y = diff_mean, fill = D_bins) ) +
  geom_bar(stat = "identity", position = position_identity(), alpha = 0.6) +
  # geom_linerange(
  #   aes(ymin = diff_mean - diff_sterr, ymax = diff_mean + diff_sterr),
  #   color="red", alpha = 0.5
  # ) +
  scale_fill_gradient2_tableau() +
  geom_point(
    aes(x = pcorr_sig, y = M_minus_S, fill = D_bins),
    shape = 21, size = 2,
    na.rm = T # to suppress the warning due to plotting only pcorr_sig cols
  ) + 
    geom_linerange(
    aes(ymin = diff_mean - diff_sterr, ymax = diff_mean + diff_sterr),
    color="red", alpha = 0.5
  ) +
  facet_wrap(~ name, scales = "free", labeller = label_wrap_gen(width=22)) +

  theme_minimal() +
  labs(
    title = "Mean Motion - Scrambled ISC parameter estimates for each Depth bin in each Juelich region across participants",
    subtitle = "Individual participants are plotted only for bins which show a (corrected) sig difference between Motion and Scrambled"
  ) +
  ylab("Mean Motion - Scrambled ISC estimate across participants") + 
  xlab("Cortical Depth bins :  1 = WM/GM border, 6 = pial") +
  theme(
    strip.text.x = element_text(size = 11), # increase size of title for each subplot
    panel.spacing.y = unit(3, "lines")      # increase space between facets
  ) +
  coord_flip(ylim = c(min(res2plot$diff_mean) - 0.05, max(res2plot$diff_mean) + 0.05))


```







